Advanced Calc Homework \large 1: 1.1: 1,3,5,7,8,11,17, 1.2: 1,4

### 2017-02-02 311/latex/HW1.tex

\section{Problem 1.1.1}
\textbf{A} False, since $1$ isn't irrational.
\textbf{B} False, since $\sqrt{2}$ is irrational, and thus $2$ cannot be inductive stepped to from $1$.
\textbf{C} False. Consider $a = \sqrt{2}$ and $b = 2 - \sqrt{2}$. Both are irrational, since the sum of a rational and irrational is irrational, but their sum is a rational.
\textbf{D} False for a similar reason: consider $\pi$ and $\frac{1}{\pi}$. Both irrational since rational and irrational numbers combine via multiplication to be irrational, but have a rational product.
\textbf{E} True. ISOAC, let $n^2$ be odd but $n$ even. Thus $\exists k\in\mathbb{N} st n=2k$, so $n^2 = 4k^2 = 2*2k^2$. This is a contradiction, since $n^2$ was supposed to be odd.\\

@@

\section{Problem 1.1.5}

@@

\textbf{Proof}

Consider set $S = \{n | n = 1 \mbox{ or for } n > 1, n, n-1 \in \mathbb{N}\}$. Note that the inductive property holds for this set, since for $s \in S$, $s+1 \in S$.

@@

\section{Problem 1.1.11a}
\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $a+b \in \mathbb{R}\ \mathbb{Q}$.

\textbf{Proof} In search of a contradiction, assume for $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, that $a+b \in \mathbb{Q}$. Then by the definition of rational numbers there exist $m,n \in \mathbb{Z}$ such that $a+b = \frac{m}{n}$. $a$ can be similarly expanded using $x,y \in \mathbb{Z}: \frac{x}{y}+b = \frac{m}{n}$. We will perform algebra and show this leads to a contradiction:

\begin{align*}
  &\frac{x}{y}+b = \frac{m}{n}\\
  &b = \frac{m}{n} - \frac{x}{y}
  &b = \frac{my}{ny} - \frac{xn}{ny}
  &b = \frac{my-xn}{ny}\\
\end{align*}

Because the product and difference of two integers is also an integer, this means b can be expressed as a fraction of two integers. This is the desired contradiction. Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $a+b \in \mathbb{R}\ \mathbb{Q}$.

\section{Problem 1.1.11b}
\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $ab \in \mathbb{R}\ \mathbb{Q}$.

\textbf{Proof} In search of a contradiction, assume for $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, that $ab \in \mathbb{Q}$. Then by the definition of rational numbers there exist $m,n \in \mathbb{Z}$ such that $ab = \frac{m}{n}$. $a$ can be similarly expanded using $x,y \in \mathbb{Z}: \frac{x}{y}b = \frac{m}{n}$. This can be rewritten as $b = \frac{my}{nx}$. Because the product of two integers is itself an integer, we have expressed b as a fraction of two integers. This is the desired contradiction. Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $ab \in \mathbb{R}\ \mathbb{Q}$.

@@

\section{Problem 1.2.4}
\textbf{A} Set $\{1/n | n \in \mathbb{N}\}$
\textit{Maximum} - 1, all $n > 1$ are smaller.\\
\textit{Minimum} - None. For any $1/n$ in the set, $1/(n+1)$ is smaller.\\
\textit{Supremum} - 1. $1/1$ is the largest value in the set.\\
\textit{Infimum} - 0. Any number $\epsilon > 0$ has a number $1\m$ smaller than it, by the Archimedean Principle, so 0 is the largest lower bound.\\

\textbf{B} Set $\{x | x \in \mathbb{R} \mbox{ and } x^2 < 2\}$
\textit{Maximum} - None. For any number $e < \sqrt{2}$, $e + \frac{\sqrt{2}-e}{2}$ is bigger
\textit{Minimum} - Same argument with $-\sqrt{2}$.\\
\textit{Supremum} - $\sqrt{2}$. Anything smaller is in the set, and it isn't in the set.
\textit{Infimum} - $-\sqrt{2}$ by the same argument.\\
\end{document}

>>>>

\section{Problem 1.1.1}
\textbf{A} False, since $1$ isn't irrational.\\
\textbf{B} False, since $\sqrt{2}$ is irrational, and thus $2$ cannot be inductive stepped to from $1$.\\
\textbf{C} False. Consider $a = \sqrt{2}$ and $b = 2 - \sqrt{2}$. Both are irrational, since the sum of a rational and irrational is irrational, but their sum is a rational.\\
\textbf{D} False for a similar reason: consider $\pi$ and $\frac{1}{\pi}$. Both irrational since rational and irrational numbers combine via multiplication to be irrational, but have a rational product.\\
\textbf{E} True. ISOAC, let $n^2$ be odd but $n$ even. Thus $\exists k\in\mathbb{N} st n=2k$, so $n^2 = 4k^2 = 2*2k^2$. This is a contradiction, since $n^2$ was supposed to be odd.\\

@@

\section{Problem 1.1.5}
Before we do this proof, we'll need to prove a subclaim:\\
\textbf{WTS} S(k): $(1 + 2 + ... + k) = \frac{k(k+1)}{2}$\\

\textbf{Proof} We will use the Principle of Induction to prove the above claim for all S(k). First, the base case: S(1): $1 = \frac{1(2)}{2} = 1$. Then the inductive step:

\begin{align*}
  (1 + 2 + ... + k) &= \frac{k(k+1)}{2}\\
  1 + 2 + ... + k + k + 1 &= \frac{k(k+1)}{2} + k + 1\\
  &= \frac{k^2 + 3k + 2}{2}\\
  &= \frac{(k+1)(k+2)}{2}\\
\end{align*}

thus $S(k) \rightarrow S(k+1)$, so S(n) is true for all $n \in \mathbb{N}$. Now to prove the problem claim.\\

\textbf{WTS} S(k): $(1 + 2 + ... + k)^2 = 1^3 + 2^3 + ... + k^3$\\

\textbf{Proof} We will use the Principle of Induction to prove that S(k) is true for all $k \in \mathbb{N}$.\\

\textbf{Base case:} $1^2 = 1^3 = 1$.\\

\textbf{Inductive step:} Let $(1 + 2 + ... + k)^2 = 1^3 + 2^3 + ... + k^3$. Then by algebra:

\begin{align*}
  (1 + 2 + ... + k + k + 1)^2 &= (1 + 2 + ... + k)(1 + 2 + ... + k + k + 1) + (k + 1)(1 + 2 + ... + k + k + 1)\\
  &= (1 + 2 + ... + k)^2 + 2(k+1)(1 + 2 + ... + k) + (k + 1)^2\\
  &= (1 + 2 + ... + k)^2 + 2(k+1)\frac{k(k+1)}{2} + (k + 1)^2\\
  &= (1 + 2 + ... + k)^2 + k(k+1)^2 + (k + 1)^2\\
  &= (1 + 2 + ... + k)^2 + (k+1)^3\\
  &= 1^3 + 2^3 + ... + k^3 + (k+1)^3\\
\end{align*}

as desired. Thus since S(1) and $S(k) \rightarrow S(k+1)$, the claim S(n) must be true for all $n \in \mathbb{N}$.\\

@@

\textbf{Proof} First we will prove that set $S = {n | n = 1 \mbox{ or } n, n-1 \in \mathbb{N}}$ is inductive. First note that $S$ holds the base case property, since $1 \in S$. Also note that if $s = n - 1$, then $s \in S \rightarrow s + 1 \in S$ by the requirement that both $n$ and $n+1$ be in S simultaneously. Thus S is an inductive set.\\

Because set $S$ is inductive, if $n \in \mathbb{N}$ and $n > 1$ then $n \in S$, and also $n-1 \in S$ by definition of membership in $S$. Thus $n > 1 \in \mathbb{N} \rightarrow n-1\in\mathbb{N}$.

@@

\section{Problem 1.1.11a}
\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\setminus \mathbb{Q}$, then $a+b \in \mathbb{R}\setminus \mathbb{Q}$.

\textbf{Proof} In search of a contradiction, assume for $a \in \mathbb{Q}$ and $b \in \mathbb{R}\setminus \mathbb{Q}$, that $a+b \in \mathbb{Q}$. Then by the definition of rational numbers there exist $m,n \in \mathbb{Z}$ such that $a+b = \frac{m}{n}$. $a$ can be similarly expanded using $x,y \in \mathbb{Z}: \frac{x}{y}+b = \frac{m}{n}$. We will perform algebra and show this leads to a contradiction:

\begin{align*}
  &\frac{x}{y}+b = \frac{m}{n}\\
  &b = \frac{m}{n} - \frac{x}{y}\\
  &b = \frac{my}{ny} - \frac{xn}{ny}\\
  &b = \frac{my-xn}{ny}\\
\end{align*}

Because the product and difference of two integers is also an integer, this means b can be expressed as a fraction of two integers. This is the desired contradiction. Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\setminus \mathbb{Q}$, then $a+b \in \mathbb{R}\setminus \mathbb{Q}$.

\section{Problem 1.1.11b}
\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\setminus \mathbb{Q}$, then $ab \in \mathbb{R}\setminus \mathbb{Q}$.

\textbf{Proof} In search of a contradiction, assume for $a \in \mathbb{Q}$ and $b \in \mathbb{R}\setminus \mathbb{Q}$, that $ab \in \mathbb{Q}$. Then by the definition of rational numbers there exist $m,n \in \mathbb{Z}$ such that $ab = \frac{m}{n}$. $a$ can be similarly expanded using $x,y \in \mathbb{Z}: \frac{x}{y}b = \frac{m}{n}$. This can be rewritten as $b = \frac{my}{nx}$. Because the product of two integers is itself an integer, we have expressed b as a fraction of two integers. This is the desired contradiction. Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\setminus \mathbb{Q}$, then $ab \in \mathbb{R}\setminus \mathbb{Q}$.

@@

\section{Problem 1.2.4}
\textbf{A} Set $\{1/n | n \in \mathbb{N}\}$\\
\textit{Maximum} - 1, all $n > 1$ are smaller.\\
\textit{Minimum} - None. For any $1/n$ in the set, $1/(n+1)$ is smaller.\\
\textit{Supremum} - 1. $1/1$ is the largest value in the set.\\
\textit{Infimum} - 0. Any number $\epsilon > 0$ has a number $1\m$ smaller than it, by the Archimedean Principle, so 0 is the largest lower bound.\\

\textbf{B} Set $\{x | x \in \mathbb{R} \mbox{ and } x^2 < 2\}$\\
\textit{Maximum} - None. For any number $e < \sqrt{2}$, $e + \frac{\sqrt{2}-e}{2}$ is bigger\\
\textit{Minimum} - Same argument with $-\sqrt{2}$.\\
\textit{Supremum} - $\sqrt{2}$. Anything smaller is in the set, and it isn't in the set.\\
\textit{Infimum} - $-\sqrt{2}$ by the same argument.\\
\end{document}
